### 2019-03-09 reply_to_reviews.R1.tex
R1

\documentclass[man]{apa6}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdftitle={Reply to Reviews},
            pdfauthor={Daniël Lakens~\& Marie Delacre},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

%%% Use protect on footnotes to avoid problems with footnotes in titles
\let\rmarkdownfootnote\footnote%
\def\footnote{\protect\rmarkdownfootnote}

  \title{Reply to Reviews}
    \author{Daniël Lakens\textsuperscript{1}~\& Marie Delacre\textsuperscript{2}}
    \date{}

\shorttitle{Reply R1}
\affiliation{
\vspace{0.5cm}
\textsuperscript{1} Eindhoven University of Technology, Eindhoven, The Netherlands\\\textsuperscript{2} Service of Analysis of the Data, Université Libre de Bruxelles, Belgium}
\usepackage{csquotes}
\usepackage{upgreek}
\captionsetup{font=singlespacing,justification=justified}

\usepackage{longtable}
\usepackage{lscape}
\usepackage{multirow}
\usepackage{tabularx}
\usepackage[flushleft]{threeparttable}
\usepackage{threeparttablex}

\newenvironment{lltable}{\begin{landscape}\begin{center}\begin{ThreePartTable}}{\end{ThreePartTable}\end{center}\end{landscape}}

\makeatletter
\newcommand\LastLTentrywidth{1em}
\newlength\longtablewidth
\setlength{\longtablewidth}{1in}
\newcommand{\getlongtablewidth}{\begingroup \ifcsname LT@\roman{LT@tables}\endcsname \global\longtablewidth=0pt \renewcommand{\LT@entry}[2]{\global\advance\longtablewidth by ##2\relax\gdef\LastLTentrywidth{##2}}\@nameuse{LT@\roman{LT@tables}} \fi \endgroup}

\DeclareDelayedFloatFlavor{ThreePartTable}{table}
\DeclareDelayedFloatFlavor{lltable}{table}
\DeclareDelayedFloatFlavor*{longtable}{table}
\makeatletter
\renewcommand{\efloat@iwrite}[1]{\immediate\expandafter\protected@write\csname efloat@post#1\endcsname{}}
\makeatother
\usepackage{lineno}

\linenumbers

\authornote{

Correspondence concerning this article should be addressed to Daniël
Lakens, Den Dolech 1, IPO 1.33, 5600 MB, Eindhoven, The Netherlands.
E-mail:
\href{mailto:D.Lakens@tue.nl}{\nolinkurl{D.Lakens@tue.nl}}}

\abstract{
We respond to all the comments raised by the two reviewers and the
editors. We would like to thank the editor and reviewers for their
suggestions for improvements, which we believe made our main argument
much clearer, focussed, and formally correct. We feel we were able to
accomodate all suggestions, and we hope you think our submission is
ready to be accepted for publication in Meta-Psychology.

}

\begin{document}
\maketitle

Dear editor,

Thank you for your comments on our submitted manuscript. We have
incorporated your three comments:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  We moved the first sentence on page 3 into the next paragraph (well
  spotted).
\item
  The code for the Shiny app was on GitHub, and the GitHub repository is
  linked to the OSF project page, but we have added a component to the
  OSF page and also directly added the Shiny App code there to make it
  easier to find: \url{https://osf.io/7f8uq/}
\item
  We now immediately refer to Figure 4 in the beginning of the paragraph
  in which the Figure is described.
\end{enumerate}

We will respond to the points raised by both reviewers in turn. Our
points below detail 1) that we limit ourselves to a comparison of SGPV
and TOST, and do not require researchers to perform both an equivalence
test and a NHST (and therefore we did not add these curves to the
Figures), 2) we rewrote the frequency claims, 3) we provide more detail
about confidence intervals around correlations, and 4) removed all
statements about the SGPV being a descriptive statistic.

\section{Reply to Reviewer 1}\label{reply-to-reviewer-1}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  We have changed the part of the sentence \enquote{Our conclusions
  about the relationship between TOST \emph{p}-values and SGPV in this
  article are not dependent upon any specific example} into \enquote{Our
  conclusions about the relationship between TOST \emph{p}-values and
  SGPV hold for an SGPV calculated from confidence intervals, and
  assuming continuous normally distributed data.}. We mainly intended to
  refer to our conclusion that: \enquote{When confidence intervals are
  symmetric we can think of the SGPV as a straight line that is directly
  related to the \emph{p}-value from an equivalence test for three
  values} namely p = 0.5 for TOST when the SGPV will be 0.5, and the
  \emph{p}-values for TOST of 0.025 and 0.975 (corresponding to the
  largest observed mean where the SGPV is 0 and the smallest observed
  mean where the SGPV is 1). We have now made it clearer that when we
  talk about \enquote{correspondence between TOST \emph{p}-values and
  the SGPV}, we mean that there is a direct relation between three
  points.
\item
  We appreciate correcting our use of the term \enquote{unrelated} and
  teaching us the word \enquote{bijective}. We used the word
  \enquote{unrelated} somewhat naievely, looking at the Figures and
  wanting to express that across the three identified situations, a
  different TOST \emph{p}-value does not imply a different SGPV - the
  SGPV across these ranges of means is aways the same (either 0, 0.5, or
  1) while the TOST \emph{p}-value differs across this range of means.
  We now just directly say what we originally meant, and no longer use
  the word \enquote{unrelated}. For example, \enquote{There are three
  situations where \emph{p}-values for TOST differentiate between
  observed results, while the SGPV does not differentiate.} and
  \enquote{While the SGPV is 1 as long as the confidence interval falls
  completely within the equivalence bounds, the \emph{p}-value for the
  TOST continues to differentiate between results as a function of how
  far the confidence interval lies within the equivalence bounds (the
  further the confidence interval is from both bounds, the lower the
  \emph{p}-value).} and similarly \enquote{A third situation in which
  the SGPV remains stable across a range of observed effects, while the
  TOST \emph{p}-value continues to differentiate, is whenever the CI is
  wider than the equivalence range, and the CI overlaps with the upper
  \emph{and} lower equivalence bound.} We have removed the word
  \enquote{unrelated} throughout the manuscript.
\item
  Please do not feel like you ever need to apologize for having formal
  training we lack, that allows you to point out whenever we make
  formally incorrect statements. This is exactly why we submit our work
  for peer review - to be able to improve it based on feedback from
  experts. We now preface the sentence with \enquote{As long as data is
  normally distributed,}. We follow your recommendation throughout to
  not generalize beyond the scenarios we study. In the second paragraph
  of the manuscript we now clearly state: \enquote{We limit our analysis
  to normally distributed continuous data.} See also point 5 below.
\item
  After making the change at point 2 above, we now think the sentence
  \enquote{The conclusion of this section on asymmetric confidence
  intervals is that a SGPV of 1 or 0 can still be interpreted as a p
  \textless{} 0:025 or p \textgreater{} 0:975 in an equivalence test,
  since the SGPV and \emph{p}-value for the TOST procedure are always
  directly related at these values.} should no longer be confusing. To
  make the sentence clearer, we have changed \enquote{these values} into
  \enquote{at the values p = 0.025 and p = 0.975} It should now be clear
  the sections are not contradictory - TOST differentiates below p
  \textless{} 0.025, when the SGPV is 1 for all p \textless{} 0.025.
\item
  We agree that is unclear whether the original statement by Blume et al
  that \enquote{the degree of overlap conveys how compatible the data
  are with the null premise} holds in general. We have now amended the
  sentence to read: \enquote{Although a SGPV of 1 or 0 can be directly
  interpreted, a SGPV between 0 and 1 is not interpretable as
  \enquote{compatibility with the null hypothesis} under the assumption
  of normally distributed data, and the generalizability of this
  statement needs to be examined beyond normally distributed data
  requires further examination.}
\item
  We have removed all our earlier statements that the SGPV only works as
  a descriptive statistic. We deleted the sentence \enquote{the SGPV is
  a descriptive statistic (unlike the \emph{p}-value that is calculated
  for an equivalence test, which is an inferential statistic) that
  provides the proportion of overlap of the confidence interval and the
  equivalence range.} and deleted \enquote{The novelty of the SGPV lies
  in its use as a descriptive statistic} from the conclusion. Indeed, we
  acknowledge that \enquote{Although a SGPV of 1 or 0 can be directly
  interpreted, a SGPV between 0 and 1 is not interpretable as
  \enquote{compatibility with the null hypothesis}.} So the SGPV is an
  inferential statistic when interpreted dichotomously. We will leave
  our uncertainty of how specific SGPV values (e.g., SGPV = 0.3) should
  be interpreted to future researchers, and we do not make any
  statements about whether such values are descriptive or inferential.
\item
  We agree that using TOST for small samples (e.g., N = 4) will not be
  useful. Indeed, an a-priori power analysis for equivalence tests would
  prevent anyone from planning such a study, since the result will be
  that such a study has 0 power - the CI will never fit withing the
  equivalence bounds.
\item
  With our recommendations, we have limited our conclusions to the
  specific cases we have examined, and noted the possibility for future
  research to examine a wider set of assumptions. An additional change
  not mentioned above is that we renamed the section on asymmetric
  confidence intervals to the more specific \enquote{The relation
  between equivalence tests and SGPV for asymmetrical confidence
  intervals around correlations} instead of the more general
  \enquote{The relation between equivalence tests and SGPV when
  confidence intervals are not symmetrical}. We appreciate the very
  positive evaluation of our manuscript, but even more so the criticism
  which allowed us to improve the manuscript.
\end{enumerate}

\section{Reply to Reviewer 2}\label{reply-to-reviewer-2}

Before listing major points, the reviewer raises the point that we
compare SGPV with \emph{p}-values from TOST throughout our paper, and in
all figures, but then at the end, in page 20, we write:

\enquote{Equivalence tests combined with null-hypothesis significance
tests also allow researchers to conclude an effect is significant and
equivalent (i.e., statistically different from zero, but also too small
to be considered meaningful). Thus, the SGPV is used to classify results
into one of three possible categories (with the data falling inside or
outside the equivalence range, or being inconclusive), while equivalence
tests combined with null-hypothesis tests classify results into four
possible categories.}

The reviewer says that \enquote{the authors suddenly switch the premise;
now the SGPV should be replaced by two \emph{p}-values, one from an
equivalence test and one from a NHST}. However, we do not suggest to
replace the SGPV with an equivalence test and a NHST at all. In our
manuscript, we already pointed out how the SGPV is basically the same as
an equivalence test and a minimum effect test (on page 7):

\enquote{The third point where the SGPV and the \emph{p}-value from the
TOST procedure should overlap is where the 95\% CI falls completely
outside of the equivalence range, but one endpoint of the confidence
interval is equal to the equivalence bound (see situation C in Figure
3), when the \emph{p}-value will always be 0.975. Note that this
situation is in essence a minimum-effect test (Murphy, Myors, \& Wolach,
2014). The goal of a minimum-effect is not just to reject a difference
of zero, but to reject the smallest effect size of interest (i.e., the
equivalence bounds). The SGPV summarizes the information from an
equivalence test and a minimum-effect test.}

One might argue that we suggest to replace the SGPV with two tests.
However, the overlap of the TOST \emph{p}-value at 0.975 when SGPV = 0
gives away that the equivalence test can be interpreted as a minimum
effect test against the equivalence bound. When the TOST \emph{p}-value
is 0.975, the 95\% CI touches either the upper or lower equivalence
bound, and fall completely outside of the equivalence range. This is
mathematically identical to a statistically significant minimum effect
test at p \textless{} 0.05. We didn't make this explicit, but we clearly
should have. We have added the sentence (on page 7):

\enquote{An equivalence test and minimum effect test against the same
equivalence bound are complementary, and when a TOST \emph{p}-value is
larger than 0.975, the \emph{p}-value for the minimum effect test is
smaller than 0.05, and the minimum effect test provides no additional
information that can not be derived from the \emph{p}-value from the
equivalence test.}

One might wonder why then, on page 20, why we introduce the idea to
perform both a null-hypothesis significance test and an equivalence
test. We were not clear enough. And indeed, the sentence \enquote{This
interpretation of a SGPV as allowing researchers to reject the null,
reject the presence of a meaningful effect, or remaining inconclusive is
very similar to the Neyman-Pearson interpretation of combining a
null-hypothesis test and an equivalence test} was simply confusing. We
have deleted this sentence and replaced it with \enquote{As already
mentioned, this interpretation of a SGPV as is very similar to the
Neyman-Pearson interpretation of an equivalence test and a minimum
effect tests (which are complementary).} Furthermore, to prevent the
apparent confusion about why we suggested to perform a null-hypothesis
test and equivalence test: part of this confusion was our original
sentence comparing the SGPV and a combination of NHST and an equivalence
test which was not needed, and we understand it made it seem as if SGPV
needs to be replaced by a combination of other tests. But that is not
what we propose. We rewrote the paragraph:

\enquote{Some researchers might be interested in combining an
equivalence tests with a null-hypothesis significance test. This allows
a researcher to ask whether there is an effect that is statistically
different from zero, and whether effect sizes that are considered
meaningful can be rejected. Equivalence tests combined with
null-hypothesis tests classify results into four possible categories,
and for example allow researchers to conclude an effect is significant
and equivalent (i.e., statistically different from zero, but also too
small to be considered meaningful, see (Lakens et al., 2018)).}

We hope this makes it clear we are not pulling a bait and switch. The
SGPV is strongly related to an equivalence test (or it's complement, a
minimum effect test). Describing the relation between the SGPV and TOST
is our main contribution. We just wanted to mention that researchers
might want to combine an equivalence test with a NHST, because in the
literature, this is a question we see many psychologists ask.

We will now discuss the major points in turn.

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  In line with the comments by reviewer 1, in the second paragraph of
  the manuscript we now clearly state: \enquote{We limit our analysis to
  normally distributed continuous data.} After all, \emph{p}-values and
  confidence are directly related, but Bayesian credible intervals would
  not necessarily be, depending on the prior. Blume et al (2018) limit
  their examples to confidence intervals, but explicitly acknowledge
  other possibilities (such as Bayesian intervals). We look forward to
  future papers that illustrate the SGPV with other interval estimates
  and are happy to add the limitation to continuous normally distributed
  data in our article. Note that we did not add that the correspondence
  only holds for situations where the SE's are small compared to the
  width of the equivalence range, as the reviewer suggests.
\item
  The reviewer repeats that the correspondence only holds when SE's are
  small compared to the width of the equivalence range. The
  correspondence is not in the shape of the curve (which clearly differs
  across Figures 2, 6, and 9). We clearly state that: \enquote{When
  confidence intervals are symmetric we can think of the SGPV as a
  straight line that is directly related to the \emph{p}-value from an
  equivalence test for three values.} This correspondence is not
  dependent upon the SE's. It will always hold for continuous normal
  data. The reviewer says that \enquote{The paper does not provide
  theoretical justification for this connection nor any proofs.} We are
  unsure which proof is needed for the overlap between the three points,
  since it is rather trivial and we believe the explanation in the paper
  is very clear.
\end{enumerate}

For the relation between TOST p = 0.5 and SGPV = 0.5, TOST p = 0.025 and
SGPV = 1, and TOST p = 0.975 and SGPV = 1: ``The SGPV is 50\% when the
observed mean falls exactly on the lower or upper equivalence bound,
because 50\% of the symmetrical confidence interval overlaps with the
equivalence range. When the observed mean equals the equivalence bound,
the difference between the mean in the data and the equivalence bound is
0, the \emph{t}-value for the equivalence test is also 0, and thus the
\emph{p}-value is 0.5 (situation A, Figure 3). Two other points always
have to overlap. When the 95\% CI falls completely inside the
equivalence region, and one endpoint of the confidence interval is
exactly equal to one of the equivalence bounds (see situation B in
Figure 3) the TOST \emph{p}-value (which relies on a one-sided test) is
always 0.025, and the SGPV is 1. The third point where the SGPV and the
\emph{p}-value from the TOST procedure should overlap is where the 95\%
CI falls completely outside of the equivalence range, but one endpoint
of the confidence interval is equal to the equivalence bound (see
situation C in Figure 3), when the \emph{p}-value will always be 0.975,
and the SGPV is 0.

We do not think we can provide any other proof for this. A confidence
interval and a \emph{p}-value are directly related. The TOST procedure
is a test using the confidence interval, and the SGPV (now that we
clearly limited our discussion to a SGPV based on a confidence interval
calculated from continuous normal data) compared against the same
equivalence bounds will be related as we describe. But we believe our
article might not have been clear enough about that we limit our
discussion about the correspondence of the curves to these 3 points.
Indeed, in some figures, the curves are less similar (e.g., Figure 6, 9,
13). Nevertheless, in these graphs, we still have the correspondence
between these 3 points (or at least the 2 points if the CI is too wide
compared to the equivalence bounds to ever observe a \emph{p} = 0.025).
To make it clearer that we talk about the overlap in these 3 critical
points, we have added:

\enquote{When we discuss the relationship between the \emph{p}-values
from TOST and the SGPV, we focus on their correspondence at three
values, namely where the TOST \emph{p} = 0.025 and SGPV is 1, where the
TOST \emph{p} = 0.5 and SGPV = 0.5, and where the TOST \emph{p} = 0.975
and SGPV = 1. These three values are important for the SGPV because they
indicate the values at which the SGPV indicates the data should be
interpreted as compatible with the null hypothesis (SGPV = 1), or with
the alternative hypothesis (SGPV = 0), or when the data are strictly
inconclusive (SGPV = 0.5). These three points of overlap are indicated
by the horizontal dotted lines in Figure 2.}

We believe this satisfactorily addresses the point by the reviewer that
\enquote{Because of this lack of discussion, the reader can easily miss
that the close correspondence only occurs when the sample size is large
enough that the standard error is very small relative to the width of
the equivalence zone.} Indeed, the correspondence between the TOST and
SGPV at these 3 points should be obvious to anyone who understands the
relationship between a \emph{p}-value and a confidence interval, so we
do not provide any further proof, but we hope that clarifying we talk
about the correspondence of these 3 points, and not the entire curve, is
sufficient to address point 2.

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{2}
\item
  We are aware of the Beger and Hsu (1996) article. But if the reviewer
  reads more closely, it should be clear that we never say the TOST is
  UMP. We say that TOST is a test procedure that aims to control error
  rates - our use of the term \enquote{Neyman-Pearson perspective}
  refers to the classic difference between significance tests (Fisher)
  and hypothesis tests (Neyman-Pearson). We do not imply that any test
  that fixes an error rate is a NP test. Because many psychologists are
  not trained in the difference between a Fisherian approach and a
  Neyman Pearson approach to the use of \emph{p}-values, we are simply
  adding a \enquote{Neyman-Pearson} approach in the text to explain that
  \enquote{From a Neyman-Pearson perspective on statistical inferences
  the main conclusion is based on whether or not \(p < \alpha\), and
  thus an equivalence test and SGPV can be performed by simply checking
  whether the confidence interval falls within the equivalence range,
  just as a null-hypothesis test can be performed by checking whether
  the confidence interval contains zero or not.} We believe this is a
  correct use of the term \enquote{Neyman-Pearson perspective on
  statistical inferences}.
\item
  We tried to find the passage the reviewer refers to when he writes
  \enquote{In light of this, the assertion without proof that classical
  \emph{p}-values provide more or better assessments of the
  \enquote{evidence} appear out of place.} but we could not. We do not
  use the word \enquote{evidence} in the text (because neither the
  \emph{p}-value from TOST nor the SGPV are measures of
  \enquote{evidence} - we fully agree with BLume et al., 2018, in this
  respect). We also do not say the \emph{p}-value is \enquote{naturally
  informative}. Indeed, related to point 3 above, we tried to be
  consistent in interpreting the TOST procedure from a Neyman-Pearson
  perspective, and focus on error control. So we regrettably do not know
  how to incorporate this point.
\item
  We fully agree with the reviewer on point 5 that \enquote{The maximum
  Type I Error rate for the SGPV over the null interval is 5\%, exactly
  like the TOST test.} Indeed, given the direct overlap between the TOST
  \emph{p}-value and the SGPV at \emph{p} = 0.025 and 0.0975 and SGPV of
  1 and 0, any error rates (when concluding support for the null or
  alternative when the SGPV is either 1 or 0) must be identical. This is
  why we were so surprised when Blume et al (2018) write \enquote{They
  control the Type I error naturally, forcing it to zero as the sample
  size grows. This, in turn, offsets Type I Error inflation that results
  from multiple comparisons or multiple examinations of accumulating
  data. Findings identified by second-generation \emph{p}-values are
  less likely to be false discoveries than findings identified by
  classical \emph{p}-values. Consequently, second-generation
  \emph{p}-values do not require ad-hoc adjustments to provide strict
  error control and this improves power in studies with massive multiple
  comparisons.} Clearly, it can not both be true that the error rate of
  SGPV and TOST is identical, TOST requires adjusting for multiple
  comparisons, but the SGPV does not.
\end{enumerate}

The editor asks us to better substantiate the frequency claims for the
comparison of TOST and SGPV, but since the SGPV and TOST \emph{p}-value
are directly related (a TOST \emph{p} \textless{} 0.025 = a SGPV of 1)
we do not think any additional substantiation is possible, especially
since the reviewer agrees that (as long as confidence intervals are used
to calculate the SGPV, which we now clearly limit our conclusions to)
\enquote{The maximum Type I Error rate for the SGPV over the null
interval is 5\%, exactly like the TOST test}.

Nevertheless, we agree this discussion was a bit too extensive. Given my
own fondness of error control in statistical inferences, I get worried
if researchers suggest we can ignore it, especially when making multiple
comparisons. Imagine people start to use the SGPV instead of the TOST
procedure whenever they perform 20 tests, and then argue they do not
need to correct for multiple comparison - I think we can all agree this
is undesireable! We have reduced the two paragraphs to one hopefully
clear statement about the need to control error rates, also when using
the SGPV. We hope this addresses the concerns by the editor as well as
the reviewer, while still making an important point in the literature
about the need to control error rates (if one cares about drawing
incorrect conclusions).

Blume, D'Agostino McGowan, Dupont, and Greevy (2018) claim that
\enquote{Adjustments for multiple comparisons are obviated} (p.~15).
However, this is not correct. Given the direct relationship between TOST
and SGPV highlighted in this manuscript (where to TOST \emph{p} = 0.025
equals SGPV = 1, as long as the SGPV is calculated based on confidence
intervals, and assuming continuous and normally distributed data), not
correcting for multiple comparisons will inflate the probability of
concluding the absence of a meaningful effect based on the SGPV in
exactly the same way as for equivalence tests. Whenever statistical
tests are interpreted as support for a hypothesis (e.g., SPGV = 0 or
SGPV =1), it is possible do so erroneously, and if researchers want to
control error rates, they need to correct for multiple comparisons.

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{5}
\item
  We have removed this section while addressing point 5 above, and thus
  incorporated the reviewers'suggestion to remove these statements.
\item
  We have not removed the paragraph about multiple comparisons, as the
  reviewer asks, but we have simplified our treatment of it. We do not
  need any other argument to make our claims about multiple comparisons
  that fact, noted by the reviewer, that \enquote{The maximum Type I
  Error rate for the SGPV over the null interval is 5\%, exactly like
  the TOST test.} This is true, and it logically follows (hence, no
  mathematical proof is needed) that error rates need to be controlled
  when using SGPV to conclude support for the null-hypothesis. We
  believe error control is important enough to include this paragraph in
  the paper, as we believe that lack of error control was one of the
  causes of the replication crisis in psychology.
\item
  We thank the reviewer for this comment. Indeed, to calculate
  confidence intervals around a correlation, the Fisher's
  z-transformation is used, which is approximately normally distributed,
  and symmatric CI are calculated. When transforming these confidence
  intervals back to correlations, they become asymmetric. We have added
  to the text: \enquote{Note that calculating confidence intervals for a
  correlation involves a Fisher's z-transformation, which transforms
  values such that they are approximately normally distributed, which
  allows one to compute symmetric confidence intervals, which are then
  retransformed into a correlation, where the confidence intervals are
  asymmetric if the correlation is not exactly zero.} This is simply one
  example of a general effect mentioned in Blume et al (2018), which we
  already cited in the main text: \enquote{The magnitude of an
  inconclusive second-generation \emph{p}-value can vary slightly when
  the effect size scale is transformed. However definitive findings,
  i.e.~a \emph{p}-value of 0 or 1 are \emph{not} affected by the scale
  changes.}
\end{enumerate}

Minor Concerns:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  We added \enquote{,or more extreme data,}
\item
  This text was deleted (see major point 5 above)
\item
  We see how \enquote{reverse the SGPV} might have been confusing. We
  did not mean to imply to reverse the null and alternative. We now just
  say that we plot 1-SGPV (which is what we incorrectly tried to express
  with \enquote{reverse}).
\item
  The issue of the SGPV needing a \enquote{correction} by setting it to
  0.5 is not related to the fact that large \emph{p}-values are not
  evidence for the null. The \emph{p}-value is behaving as it is
  intended to, without any correction. A high \emph{p}-value is not a
  \enquote{misleading result} - although it can be misinterpreted (which
  is exactly what we try to prevent with our work on equivalence
  testing). We therefore did not incorporate this point.
\item
  In line with comments by reviewer 1, we removed statemetns about SGPV
  not being inferential, thereby solving the issue reviewer 2 raised.
\item
  In line with major point 8, we added information that asymmetric
  confidence intervals around a correlation are symmetric when Fisher
  z-transformed. This is just an example of what Blume et al (2018)
  refer to when they write \enquote{The magnitude of an inconclusive
  second-generation \emph{p}-value can vary slightly when the effect
  size scale is transformed. However definitive findings, i.e.~a
  \emph{p}-value of 0 or 1 are \emph{not} affected by the scale
  changes.} We believe this section, although the most detailed, is
  interesting, even only for more expert readers (indeed, Reviewer 1
  notes that \enquote{The asymmetric CI case is very interesting}). We
  have made our point less general, now only talking about correlations,
  by changing the title of this section into \enquote{The relation
  between equivalence tests and SGPV for asymmetrical confidence
  intervals around correlations}. The reviewer is correct confidence
  intervals can be created in many difference ways (e.g., with holes in
  the middle), but for this paper, we just focus on one specific case
  now, which we believe might be closest to what researchers would use
  in real life, the confidence interval around a correlation.
\item
  It is a well-known fact that the TOST can have 0 power (and indeed,
  this is one of the main issues in the Perlman and Wu (1999) paper the
  reviewer cited). As explained in Lakens (2017): \enquote{It is
  important to take statistical power into account when determining the
  equivalence bounds because, in small samples (where CIs are wide), a
  study might have no statistical power (i.e., the CI will always be so
  wide that it is necessarily wider than the equivalence bounds).} We
  agree with the reviewer it is not interesting to do a test if you have
  0 power. People should design informative experiments by doing an
  a-priori power analysis. Any problem in this respect is not a problem
  with equivalence tests, but a problem with designing a bad study. Note
  that the exact same issue applies to the SGPV. It is also possible to
  have 0 power for the SGPV - so we do not understand why the reviewer
  is so surprised about this. The question \enquote{Why use the TOST
  test then?} equally applies to \enquote{Why use the SGPV then?} if
  you, from the outset, know that the \enquote{small sample correction}
  will lead to a SGPV of 0.5.
\item
  We do not see how setting the SGPV to 0.5 \enquote{shrinks} an
  estimate to some value. Regardless of how correction factors are
  applied in statistics, we are trying to educate our readers about what
  the \enquote{small sample correction} does. We believe our readers
  will associate \enquote{correction} with \enquote{correcting}, and
  want to explain that the SGPV is not set to a \enquote{correct} value.
  We are not talking about statistical terms here - we are warning the
  reader to not interpret \enquote{correction} in terms of normal
  language use. This is similar to warning readers that
  \enquote{significant} does not have the normal language connotation of
  \enquote{important}. We have therefore kept our original sentence.
\end{enumerate}

As a final comment, altough not explicitly listed as a major point of
minor point, we are happy to respond to the reviewers' suggestion that
Perlman and Wu (1999) call for an \enquote{outright ban of TOST}. The
best summary of this discussion (of which Perlman \& Wu is one
contribution) can be found in Meyners (2012), which we cited. The
discussion has different arguments, and Meyners draws the conclusion
that there is no universially preferred approach. The power benefits of
alternative approaches are typically modest, or only noticeable in
situations where all alternatives are underpowered. The power issue is
not that important - the main discussion is about whether equivalence
tests should have a bounded rejection region, like TOST, or an unbounded
rejection region. It might be useful to note that TOST and SGPV are
similar in this respect - both have bounded rejection regions, and if
the reviewer agrees with Perlman and Wu (1999) in the criticism against
TOST, the same criticism applies to the SGPV. If the reviewer reads
Perlman and Wu (1999) as an argument to ban TOST, it is at the same time
an argument to ban the SGPV. We agree with Meyners, who writes:
\enquote{The debate has been quite heated in times; this section intends
to give a brief overview of the debate. We would like to emphasize again
that the differences between the approaches are minor and to a large
extent of academic interest only, as is this debate.} and concludes:
\enquote{To summarize, there is no right or wrong in most of the
arguments; the two opposing parties rather apply different criteria to
select the most appropriate test, one group focusing rather on
mathematical/statistical criteria, the other one more on
interpretability and reasonability.} We have added a reference to this
discussion for the interested reader, and added \enquote{For an
excellent discussion of the strengths and weaknesses of different
frequentist equivalence tests, including alternatives to the TOST
procedure, see Meyners, 2012.}

\begingroup
\setlength{\parindent}{-0.5in} \setlength{\leftskip}{0.5in}

\hypertarget{refs}{}
\hypertarget{ref-blume_second-generation_2018}{}
Blume, J. D., D'Agostino McGowan, L., Dupont, W. D., \& Greevy, R. A.
(2018). Second-generation p-values: Improved rigor, reproducibility, \&
transparency in statistical analyses. \emph{PLOS ONE}, \emph{13}(3),
e0188299.
doi:\href{https://doi.org/10.1371/journal.pone.0188299}{10.1371/journal.pone.0188299}

\endgroup

\end{document}